---
title: "MyRide | Interval Analyzer"
output: 
  flexdashboard::flex_dashboard:
    vertical_layout: fill
    orientation: rows
    horizontal_layout: scroll
    source_code: embed
    logo: media/new_logo.png
    theme: lumen
    navbar:
      - { icon: "fa-github", href: "https://github.com/brndngrhm/my_ride", align: right}
runtime: shiny
---

<style>   

.navbar {
background-color:#272822;
border-color:#f92672;
}

</style> 

```{r setup, include=FALSE}
library(dplyr)
library(lubridate)
library(stringr)
library(tidyr)
library(changepoint)
library(zoo)
library(flexdashboard)
library(highcharter)
library(shiny)
library(leaflet)
library(XML) 
library(xml2)
library(tidyr)
library(FITfileR)
```

Ride Data 
======================================================================

Sidebar {.sidebar}
----------------------------

```{r sidebar}

#increase max file size allowed 
options(shiny.maxRequestSize=30*1024^2) 

fileInput(inputId = "file_upload", 
          label = h3("Next, upload a FIT file and wait a little..."),
          multiple = FALSE,
          accept = c(".fit"),
          placeholder = "feed me a .fit file")

radioButtons(inputId = "measure_select",
             label = h3("Then choose what you want to see"), 
             choices = c("heart_rate", "power", "cadence", "speed"),
             selected = "heart_rate")

sliderInput(inputId = "slider_select",
            label = h3("If the blue line isn't lining up with your intervals, adjust the slider (See 'About' Tab for more info)"),
            min = 0, 
            max = 35,
            value = 15, 
            step = 1)

```

----------------------------

[How to download a .tcx file from Garmin Connect](https://support.strava.com/hc/en-us/articles/216917807-Exporting-files-from-Garmin-Connect)

[How to download a .tcx file from Strava](https://support.strava.com/hc/en-us/articles/216918437-Exporting-your-Data-and-Bulk-Export#TCX)

```{r}
#sometimes data is read in as multiple records, fomrming a list. in that case we need to bind them together to 1 dataframe
parse_fit_file <- function(data){
  
  fit_file <- FITfileR::readFitFile(data)
  
  fit_file_df <- if(class(records(fit_file)) %in% "list"){
    
    purrr::map_dfr(records(fit_file), bind_rows)
    
  } else {
    
    records(fit_file)
    
  }
  
  return(fit_file_df)
  
}

# minute averages of all numeric variables
get_minute_avgs <- function(data){
  
  fit_file_formatted <- data %>%
    as_tibble() %>%
    mutate(timestamp = ymd_hms(timestamp) - hours(5), 
           time = floor_date(timestamp, "minute")) %>%
    select(-accumulated_power) %>%
    group_by(time) %>%
    summarise(across(where(is.numeric), mean, na.rm = T))%>%
    mutate(count = 1,
           ride_min = cumsum(count)
    )
  
  return(fit_file_formatted)
  
}

# lets you select 1 variable at a time 
get_specific_value <- function(data, value){
  
  vars <- c("time", "ride_min", {{value}})
  
  data %>%
    select(one_of(vars)) %>% 
    rename(metric_avg = 3)

}

# does changepoint analysis on a variable
get_changepoints <- function(data, sensitivity){
  
#changepoint analysis - http://members.cbio.mines-paristech.fr/~thocking/change-tutorial/RK-CptWorkshop.html
cp_mean <- cpt.mean(data$metric_avg,
                    test.stat = 'Normal',
                    method = 'BinSeg',
                    # Q=input$slider_select,
                    Q = sensitivity,
                    penalty = "SIC")

#store breakpoints in a dataframe
changepoints <- data.frame("ride_min" = cpts(cp_mean),
                           changes = seq(1, length(cpts(cp_mean)), 1))

#join breakpoints to formatted data, fill in the gaps using Zoo::na.locf
data <- left_join(data, changepoints, by = "ride_min") %>%
  tidyr::fill(changes, .direction = "downup") 

#calculate the mean within each segment, as identified by the changepoints
changepoint_mean <- data %>%
  group_by(changes) %>%
  summarise(mean = mean(metric_avg))

#join mean back to dataframe
left_join(data, changepoint_mean, by = "changes")

}

###################

parsed_fit_file <- reactive({
  
  infile <- input$file_upload
  
  parse_fit_file(infile$datapath)
  
})

ride_data <- reactive({
  
  validate(need(input$file_upload != "", "Upload a .tcx file. Then wait..."))

  get_minute_avgs(parsed_fit_file())
  
})

selected_value <- reactive({
  get_specific_value(ride_data(),  input$measure_select)
})

changes <- reactive({
  get_changepoints(selected_value(), sensitivity = input$slider_select)
  })

```

Row 
------------
### **Interval Analyzer**

```{r main_plot}
renderHighchart({
  hchart(changes(), "line", hcaes(x = datetime_to_timestamp(time), y=round(metric_avg,0)), color = "#A6E22E",showInLegend = F, name = input$measure_select) %>%
    hc_add_series(changes(), "line", hcaes(x = datetime_to_timestamp(time), y=round(mean)), color = "#1E90FF",  showInLegend = T, name = "Region Average") %>%
    hc_add_theme(hc_theme_monokai()) %>%
    hc_tooltip(shared = T) %>%
    hc_title(text = input$measure_select,
             align = "left") %>%
    hc_subtitle(text = if(input$measure_select == "Distance"){paste("Total Distance:", max(round(changes()$avg,0)), sep=" ")}
                         else{paste("Overall Mean:", round(mean(changes()$avg),2), sep=" ")},
             align = "left") %>%
    hc_xAxis(title = list(text = "Minutes into Ride"))%>%
    # hc_yAxis(title = list(text = unique(changes()$y_label))) %>%
    hc_chart(zoomType = "x")
})
```

### **Overlay**
```{r overlay_plot}
renderHighchart({
  
  my_vars <- c("time", "cadence", "heart_rate", "power", "speed")
  
  ride_data() %>%
    select(one_of(my_vars)) %>%
    pivot_longer(cols = 2:ncol(.)) %>%
    hchart(., "line", hcaes(x = datetime_to_timestamp(time), y = round(value,0), group = name), showInLegend = T) %>%
    hc_add_theme(hc_theme_monokai()) %>%
    hc_tooltip(shared = T) %>%
    hc_title(text = "Comparison",
             align = "left") %>%
    hc_xAxis(title = list(text = "Minutes into Ride"))%>%
    hc_yAxis(title = list(text = " ")) %>%
    hc_chart(zoomType = "x")
})

```

Row
-------------

### Distribution
```{r histogram}
#histogram for selected measure
renderHighchart({
  
    validate(
      need(sum(selected_value()$metric_avg, na.rm = T) > 0, 
           paste("There's no", input$measure_select, "data to plot :( Please select another option on the sidebar", sep=" "))
    )

  hchart(selected_value()$metric_avg) %>%
      hc_add_theme(hc_theme_monokai()) %>%
      hc_title(text = input$measure_select,
               align = "left") %>%
      hc_xAxis(title = list(text =input$measure_select))  %>%
      hc_legend(enabled = FALSE)
  
})
```

### Map
```{r map}

#ride map
renderLeaflet({
  
   validate(
      need(sum(ride_data()$position_lat, na.rm = T) > 0, 
           "There's no GPS data to plot :(")
    )
  
  coords <- ride_data() %>%
    select(starts_with("position"))
  
  leaflet(coords) %>%  
    addProviderTiles(providers$CartoDB.DarkMatter) %>%
    setView(lng = mean(coords$position_long, na.rm = T), 
            lat = mean(coords$position_lat, na.rm = T), 
            zoom = 11) %>% 
    leaflet::addPolylines(~position_long, 
                          ~position_lat,
                          weight = 4, 
                          color = "#A6E22E",
                          stroke = TRUE, 
                          fillOpacity = 0)
})

```

HR Analysis
======================================================================

Sidebar {.sidebar}
----------------------------
```{r hr_sidebar}

#increase max file size allowed 
options(shiny.maxRequestSize=30*1024^2) 

sliderInput(inputId = "hr_slider_select",
            label = h3("Adjust Lag"),
            min = 0, 
            max = 35,
            value = 15, 
            step = 1)

```

Row
--------------------

### Cross Corellation Between Heart Rate and Power
```{r ccf_plot}

#scatterplot of HR vs power

renderPlot({
  
    validate(
      need(sum(hist_data()$value, na.rm = T) > 0, paste("There's no data to plot!", sep=" "))
    )
  
  hr_data <- rides_data() %>%
    dplyr::group_by(watts) %>%
    dplyr::summarise(avg_hr = mean(heart_rate, na.rm = T))
  
  #CCF analysis (#interp: HR lags watts most sig at ~ 15 seconds)
  ccf <- ccf(subset(rides_data()$watts,!is.na(rides_data()$heart_rate)), !is.na(rides_data()$heart_rate),
             plot = T) 
  ccf.df <- data.frame(lag = ccf$lag,
                       acf = ccf$acf)
  
  #hchart(ccf)
  
  #CCF plot
  # hchart(ccf.df, type = "column", name = "ACF", hcaes(x = lag, y = round(acf, 2))) %>%
  #   hc_add_theme(hc_theme_darkunica()) %>%
  #   hc_tooltip(shared = T) %>%
  #   hc_title(text = "Cross Corellation Between Watts and Heart Rate",
  #            align = "left") %>%
  #   hc_subtitle(text = paste("Interpretation: On this ride, your heart rate lagged power output by about", ccf$lag[ccf$acf == max(ccf$acf)], "seconds", sep = " "),
  #               align = 'left') %>%
  #   hc_xAxis(title = list(text = "Lag (in Seconds)"))%>%
  #   hc_yAxis(title = list(text = "ACF")) %>%
  #   hc_chart(zoomType = "x")
  
})
```

### Lagged Watts
```{r lagged_plot}

renderHighchart({
  
    validate(
      need(sum(hist_data()$value, na.rm = T) > 0, paste("There's no data to plot!", sep=" "))
    )
  
#Shift data
lag <- rides_data() %>%
  dplyr::mutate(watts_lag = dplyr::lag(watts, n = input$hr_slider_select, default = NA)) %>%
  dplyr::group_by(watts_lag) %>%
  dplyr::summarise(avg_lag_hr = mean(heart_rate, na.rm = T))

#LM on lagged data    
lm_test <- glm(avg_lag_hr ~ watts_lag + I(watts_lag^2) + I(watts_lag^3), data = lag)
  #summary(lm_test)
  # plot(lm_test)
  
  
#added model fitted data to dataframe
estimates <- broom::tidy(lm_test)

lag$fitted <- estimates$estimate[estimates$term == "(Intercept)"] + 
  estimates$estimate[estimates$term == "watts_lag"]*lag$watts_lag + 
  estimates$estimate[estimates$term == "I(watts_lag^2)"]*lag$watts_lag^2+ 
  estimates$estimate[estimates$term == "I(watts_lag^3)"]*lag$watts_lag^3 

#plagged plot with fit line
hchart(lag, type = "line", name = "Lagged HR", hcaes(x=watts_lag, y=round(avg_lag_hr,2))) %>%
  hc_add_series(lag, name = "model fit", type = "line", hcaes(x=watts_lag, y=round(fitted,2))) %>%
  hc_add_theme(hc_theme_monokai()) %>%
  hc_tooltip(shared = T) %>%
  hc_title(text = "Lagged Watts vs Heart Rate",
           align = "left") %>%
  hc_subtitle(text = paste("Lag =", input$hr_slider_select, "Seconds", sep = " "),
           align = "left") %>%
  hc_xAxis(title = list(text = "Watts"))%>%
  hc_yAxis(title = list(text = "BPM")) %>%
  hc_chart(zoomType = "x")
  
})

```


## About
======================================================================
<style>

  .about {
    font-family: font-family: Arial, Helvetica, sans-serif;
    font-size: 18px;
    
  }

</style>

### **About** {.about}

#### **What is this app for? Why should I use this instead of something like Strava, Training Peaks, or Garmin Connect?**  
The free versions of these apps aren't very helpful for analyzing intervals. While they can tell you your overall average power, heart rate or cadence for a ride it's harder if not impossible to get at the average *during an interval or hard effort*, which is probably what you care about more so than the overall average. **This app automatically identifies interval efforts and displays the average power, heart rate, cadence, or speed in these intervals.**
<br>  

#### **How do I use the chart? What are the green and blue lines telling me?**  
Let's say you want to analyze your power during a recent training ride. You upload a file, and select "Power" and a chart appears. The Green Line is showing you your average power at each minute of your ride. The Blue Line is detects changes in your effort and tells you the average power within each detected change point. It is a result of a [changepoint detection algorithm](http://members.cbio.mines-paristech.fr/~thocking/change-tutorial/RK-CptWorkshop.html), the sensitivity of which you can control with the slider in the sidebar.
<br>  

#### **Ok but what if the Blue Line isn't lining up with my intervals?**
The Blue Line tries to detect changes automatically, but can sometimes miss them. The good news is you can adjust the number of changepoints to search for with the slider - the lower the number the less sensitive and the higher the number the more sensitive it becomes. The goal is to strike a balance, where the blue line changes in step with your your intervals but doesn't change *too much* to the point where it reacts to every small fluctuation within an interval. Usually setting it somewhere between 10 and 20 should do the trick.
<br>  

#### **Why do I need to specify if I used a power meter or not??**
The .tcx files are structured differently when you use a power meter than when you don't. You have to select the Yes or No so the app processes the data correctly. 
<br>  

#### **Why do I need to upload a file at all? Why can't this connect to Strava or Garmin Connect??**
This might be possible in the future! Currently I don't know how to extract anything other than ride summary statistics from Strava and Garmin Connect's API has an insane fee of $5,000.

[How to download a .tcx file from Garmin Connect](https://support.strava.com/hc/en-us/articles/216917807-Exporting-files-from-Garmin-Connect)

[How to download a .tcx file from Strava](https://support.strava.com/hc/en-us/articles/216918437-Exporting-your-Data-and-Bulk-Export#TCX)


