---
title: "MyRide | Interval Analyzer"
output: 
  flexdashboard::flex_dashboard:
    vertical_layout: fill
    orientation: rows
    horizontal_layout: scroll
    source_code: embed
    logo: media/new_logo.png
    theme: bootstrap
    navbar:
      - { icon: "fa-github", href: "https://github.com/brndngrhm/my_ride", align: right}
    css:
      # - bootstrap_min.css
      - dark.css
runtime: shiny
---

```{r setup, include=FALSE}
library(dplyr)
library(lubridate)
library(stringr)
library(tidyr)
library(changepoint)
library(zoo)
library(flexdashboard)
library(highcharter)
library(shiny)
library(leaflet)
library(XML) 
library(xml2)
library(tidyr)
library(FITfileR)
library(imputeTS)
library(slider)
```

Analysis
======================================================================

Sidebar {.sidebar}
----------------------------

```{r sidebar}

#increase max file size allowed 
options(shiny.maxRequestSize=30*1024^2) 

fileInput(inputId = "file_upload", 
          label = h4("Upload a FIT file"),
          multiple = FALSE,
          accept = c(".fit"),
          placeholder = ".fit format only")

actionButton(inputId = "go", label = "Analyze", width = '100%', icon = icon("fa-chart-line"))

shiny::hr()

radioButtons(inputId = "measure_select",
             label = h4("Choose what you want to see"), 
             choices = c("heart_rate", "power", "cadence", "speed"),
             selected = "power")

sliderInput(inputId = "slider_select",
            label = h4("If the blue line isn't lining up with your intervals, adjust the slider (See 'About' Tab for more info)"),
            min = 0, 
            max = 35,
            value = 15, 
            step = 1)
```

----------------------------

```{r}
#sometimes data is read in as multiple records, fomrming a list. in that case we need to bind them together to 1 dataframe
parse_fit_file <- function(data){
  
  fit_file <- FITfileR::readFitFile(data)
  
  fit_file_df <- if(class(records(fit_file)) %in% "list"){
    
    purrr::map_dfr(records(fit_file), bind_rows)
    
  } else {
    
    records(fit_file)
    
  }
  
  # checks if col names exist, and if not create it with value 0.
  expected_cols <- c("timestamp", "position_lat", "position_long", "distance", "accumulated_power", 
                     "altitude", "speed", "power", "heart_rate", "cadence", "temperature", 
                     "fractional_cadence")
  
  fit_file_df[expected_cols[!(expected_cols %in% colnames(fit_file_df))]] <- 0
  
  fit_file_df %>%
    mutate(heart_rate = ifelse(heart_rate == 255, 0, heart_rate),
           cadence_raw = cadence,
           cadence = imputeTS::na_ma(cadence_raw, k = 4, weighting = 'simple'),
           speed_raw = speed,
           speed = imputeTS::na_ma(speed_raw, k = 4, weighting = 'simple')* 2.23694,
           power_raw = power,
           power_adj = ifelse(power_raw > 1500, NA_real_, power_raw),
           power = imputeTS::na_ma(power_adj, k = 4, weighting = 'simple')
    )
  
}

# minute averages of all numeric variables
get_minute_avgs <- function(data){
  
  fit_file_formatted <- data %>%
    as_tibble() %>%
    mutate(timestamp = ymd_hms(timestamp) - hours(5), 
           time = floor_date(timestamp, "1 second")) %>%
    select(-accumulated_power) %>%
    group_by(time) %>%
    summarise(across(where(is.numeric), ~mean(.x, na.rm = T))) %>%
    mutate(count = 1,
           ride_min = cumsum(count)
    )
  
  return(fit_file_formatted)
  
}

# lets you select 1 variable at a time 
get_specific_value <- function(data, value){
  
  vars <- c("time", "ride_min", {{value}})
  
  data %>%
    select(one_of(vars)) %>% 
    rename(metric_avg = 3)
}

# does changepoint analysis on a variable
get_changepoints <- function(data, sensitivity){
  
  #changepoint analysis - http://members.cbio.mines-paristech.fr/~thocking/change-tutorial/RK-CptWorkshop.html
  cp_mean <- cpt.mean(data$metric_avg,
                      test.stat = 'Normal',
                      method = 'BinSeg',
                      Q = input$slider_select,
                      penalty = "SIC")
  
  #store breakpoints in a dataframe
  changepoints <- data.frame("ride_min" = cpts(cp_mean),
                             changes = seq(1, length(cpts(cp_mean)), 1))
  
  #join breakpoints to formatted data, fill in the gaps
  data <- left_join(data, changepoints, by = "ride_min") %>%
    tidyr::fill(changes, .direction = "downup") 
  
  #calculate the mean within each segment, as identified by the changepoints
  changepoint_mean <- data %>%
    group_by(changes) %>%
    summarise(mean = mean(metric_avg, na.rm = T))
  
  #join mean back to dataframe
  left_join(data, changepoint_mean, by = "changes")
  
}

get_normalized_power <- function(data){
  
  # calc based on: https://medium.com/critical-powers/formulas-from-training-and-racing-with-a-power-meter-2a295c661b46
  
  data %>%
    mutate(window = floor_date(time, "30 seconds"), 
           rolling_30s_avg = slide_index_dbl(metric_avg, window, mean, .before = seconds(30)),
           rolling_avg_powered = rolling_30s_avg^4) %>%
    ungroup() %>%
    mutate(avg_powered_values = mean(rolling_avg_powered),
           NP = avg_powered_values^0.25) %>%
    pull(NP) %>%
    unique()
}

###################

parsed_fit_file <- reactive({
  
  infile <- input$file_upload
  
  parse_fit_file(infile$datapath)
  
})

ride_data <- reactive({
  
  get_minute_avgs(parsed_fit_file())
  
})

selected_value <- reactive({
  get_specific_value(ride_data(),  input$measure_select)
})

changes <- reactive({
  get_changepoints(selected_value(), sensitivity = input$slider_select)
})

color <- reactive({
  
  colors <- tibble(value = c("cadence", "heart_rate", "power", "speed"),
                   color = c("#66D9EF", "#F92672", '#A6E22E', '#AE81FF')
  ) %>%
    filter(value == input$measure_select) %>%
    pull(color)
  
})

normalized_power <- reactive({
  get_normalized_power(
    get_specific_value(ride_data(), "power"))
})

```

Row 
------------
### 
```{r}

renderHighchart({
  
  validate(need(input$file_upload != "","
- Upload a FIT file.
- The app will display power, heart rate, cadence, or speed of your ride. 
- If any intervals are detected, this app will also display the average power, heart rate, cadence, or speed in these intervals.
- Large files take longer to process.
"))
  
  validate(
    need(sum(selected_value()$metric_avg, na.rm = T) > 0, 
         paste("There's no", input$measure_select, "data to plot :( Please select another option on the sidebar", sep=" "))
  )
  
  normalized_power_label <- if(input$measure_select == "power"){normalized_power()} else {"Not Applicable"}
  
  subtitle <- if(input$measure_select == "power"){
    paste("Overall Avg:", round(mean(changes()$metric_avg),2), " | Normalized:", round(mean(normalized_power_label),2))
  } else {paste("Overall Avg:", round(mean(changes()$metric_avg),2))}
  
  hchart(changes(), "line", hcaes(x = datetime_to_timestamp(time), y=round(metric_avg,0)), color = color(), showInLegend = F, name = input$measure_select) %>%
    hc_add_series(changes(), "line", hcaes(x = datetime_to_timestamp(time), y=round(mean)),
                  color = "#1E90FF", showInLegend = T, name = "Region Average") %>%
    hc_add_theme(hc_theme_monokai()) %>%
    hc_tooltip(shared = T) %>%
    hc_title(text = paste("Changepoint Analysis:", input$measure_select),
             align = "left") %>%
    hc_subtitle(text = subtitle,
                align = "left") %>%
    hc_xAxis(title = list(text = "time"),
             type = "datetime"
    ) %>% 
    hc_chart(zoomType = "x")
})
```

###
```{r}
renderHighchart({
  
  validate(need(input$file_upload != "", " "))
  
  my_vars <- c("time", "cadence", "heart_rate", "power", "speed")
  
  cols <- c("#66D9EF", "#F92672", '#A6E22E', '#AE81FF')
  
  ride_data() %>%
    select(one_of(my_vars)) %>%
    pivot_longer(cols = 2:ncol(.)) %>%
    hchart(., "line", hcaes(x = datetime_to_timestamp(time), y = round(value,0), group = name), showInLegend = T) %>%
    hc_colors(cols) %>%
    hc_add_theme(hc_theme_monokai()) %>%
    hc_tooltip(shared = T) %>%
    hc_title(text = "Data Overlay",
             align = "left") %>%
    hc_xAxis(title = list(text = "time"),
             type = "datetime"
    ) %>% 
    hc_yAxis(title = list(text = " ")) %>%
    hc_chart(zoomType = "x")
})
```

Row
----------

### 
```{r}
renderHighchart({
  
  validate(need(input$file_upload != "", " "))
  
  validate(
    need(sum(selected_value()$metric_avg, na.rm = T) > 0, 
         paste("There's no", input$measure_select, "data to plot :( Please select another option on the sidebar", sep=" "))
  )
  
  hchart(selected_value()$metric_avg, color = color()) %>%
    hc_add_theme(hc_theme_monokai()) %>%
    hc_title(text = input$measure_select,
             align = "left") %>%
    hc_xAxis(title = list(text =input$measure_select))  %>%
    hc_legend(enabled = FALSE)
  
})
```

###
```{r}
renderLeaflet({
  
  validate(need(input$file_upload != "", " "))
  
  validate(
    need(sum(ride_data()$position_lat, na.rm = T) > 0, 
         "There's no GPS data to plot :(")
  )
  
  coords <- ride_data() %>%
    select(starts_with("position"))
  
  leaflet(coords) %>%  
    addProviderTiles(providers$CartoDB.DarkMatter) %>%
    setView(lng = mean(coords$position_long, na.rm = T), 
            lat = mean(coords$position_lat, na.rm = T), 
            zoom = 11) %>% 
    leaflet::addPolylines(~position_long, 
                          ~position_lat,
                          weight = 4, 
                          color = "#A6E22E",
                          stroke = TRUE, 
                          fillOpacity = 0)
})
```

About
======================================================================
<style>

.about {
font-family: font-family: Arial, Helvetica, sans-serif;
font-size: 18px;

}

</style>

### **What is this app for? Why should I use this instead of something like Strava, Training Peaks, or Garmin Connect?**
The free versions of these apps aren't very helpful for analyzing intervals. While they can tell you your overall average power, heart rate or cadence for a ride it's harder if not impossible to get at the average *during an interval or hard effort*, which is probably what you care about more so than the overall average. **This app automatically identifies interval efforts and displays the average power, heart rate, cadence, or speed in these intervals.**
<br>  

### **How do I use the chart? What are the green and blue lines telling me?**
Let's say you want to analyze your power during a recent training ride. You upload a file, and select "Power" and a chart appears. The Green Line is showing you your average power at each minute of your ride. The Blue Line is detects changes in your effort and tells you the average power within each detected change point. It is a result of a [changepoint detection algorithm](http://members.cbio.mines-paristech.fr/~thocking/change-tutorial/RK-CptWorkshop.html), the sensitivity of which you can control with the slider in the sidebar.
<br>  

### **Ok but what if the Blue Line isn't lining up with my intervals?**
The Blue Line tries to detect changes automatically, but can sometimes miss them. The good news is you can adjust the number of changepoints to search for with the slider - the lower the number the less sensitive and the higher the number the more sensitive it becomes. The goal is to strike a balance, where the blue line changes in step with your your intervals but doesn't change *too much* to the point where it reacts to every small fluctuation within an interval. Usually setting it somewhere between 10 and 20 should do the trick.
<br>  

### **Why do I need to upload a file at all? Why can't this connect to Strava or Garmin Connect?**
This might be possible in the future! Currently I don't know how to extract anything other than ride summary statistics from Strava and Garmin Connect's API has an insane fee of $5,000.
